### 2014-11-12 tex/precise.tex
updates to sec 3.1

\section{Generating Symbolic Heap Summaries}
\label{sec:precise}

In this section we present the rules for the $\rsym$ relation. The are
three sets of rewrite rules: the initialization of symbolic
references, the field access and write rules, and finally checking the
equality and inequality of references. Here we describe each of the
rules. They consist the core of the symbolic heap summary approach.

@@

In this section we present the rewrite rules for the initialization of
references in the $\rsym$ relation for a symbolic heap summary. The
initialization rules are defined on the bi-partite graph consisting of
references and locations. In the $\rgse$ relation during
initialization of a symbolic references, it is non-deterministically
assigned to null, new instance of the object, and alias any previously
initialized symbolic reference. In contrast, the heap summarization
approach creates a single branch that contains the summarization for
all the initialization in a single bi-partitate graph.

@@

The initialization rules are invoked when an uninitialized field in a
symbolic reference is accessed. The function $\mathbb{UN}(\cfgnt{L},
\cfgnt{R}, \cfgnt{r}, \cfgnt{f}) = \{\cfgnt{l}\ ...\}$ returns
constraint-location pairs in which the field $\cfgnt{f}$ is
uninitialized:
\[
\begin{array}{rcl}
\mathbb{UN}(\cfgnt{L}, \cfgnt{R}, \cfgnt{r}, \cfgnt{f}) & = &\{ \lp\phi\ \cfgnt{l}\rp \mid \lp \phi\ \cfgnt{l}\rp  \in \cfgnt{L}\lp \cfgnt{r}\rp  \wedge \\
& & \ \ \ \ \exists \phi^\prime \lp \lp \phi^\prime\ \cfgnt{l}_\mathit{un}\rp  \in \cfgnt{L}\lp \cfgnt{R}\lp l,\cfgnt{f}\rp\rp \wedge \\
& & \ \ \ \ \ \ \ \ \mathbb{S}\lp \phi \wedge \phi^\prime \rp\rp\}\\
\end{array}
\]
where $\mathbb{S}(\phi)$ returns true if $\phi$ is
satisfiable. Intutively, for the reference, $\cfgnt{r}$, it constructs
the set, $\theta$, that contains all contraint-location pairs that
point to the field $\cfgnt{f}$ and $\cfgnt{f}$ points to
$\cfgnt{l}_\mathit{un}$.

@@

The rewrite rule to initialize a heap in the symbolic summarization
technique is shown in~\figref{fig:symInit}. Given the unitialized set
$\mathbb{UN}$ for field $f$, the $\mathit{min}_l$ function returns
$(\phi_x\ \cfgnt{l}_x)$ based on a lexical ordering of uninitialized locations
in $\Lambda$. The init$_r$ method returns a new input reference $\cfgnt{r}_f$
while the fresh$_f$ returns a new location of type $C$. The set $\rho$
contains the set of reference location pairs, $(\cfgnt{r}_a\ \cfgnt{l}_a)$, that the
field being initialized may alias. The pair $(\cfgnt{r}_a\ \cfgnt{l}_a)$ is contained
in the set $ \rho $ if the following conditions hold: (i) $\cfgnt{r}_a$ is
part of the input references initialized in the summary heap, (ii)
$\cfgnt{r}_a$ is in the pre-image of $\cfgnt{l}_a$, (iii) and the type of $\cfgnt{l}_a$ is the
same as the type of the field being initialized.

There are three constraint-location pair sets that correspond to
non-deterministic choices in GSE: (a) $\theta_\mathit{null}$ contains
a constraint $\cfgnt{r}_f = \cfgnt{r}_\mathit{null}$ paired with the location
$\cfgnt{l}_\mathit{null}$, (b) $\theta_\mathit{new}$ creates a pair with the
constraint $\cfgnt{r}_f$ is not null and it does not alias any reference in
set $\rho$ under and a newly created location $\cfgnt{l}_f$, (c) and
$\theta_\mathit{alias}$ where for each element in the set $(\cfgnt{r}_a\ \cfgnt{l}_a)
\in \rho$ it creates a constraints that $\cfgnt{r}_f$ is not null, points to
$\cfgnt{r}_a$, and does not point to any reference contained in $(\cfgnt{r}_a^\prime\ 
\cfgnt{l}_a^\prime) \in \rho$ such that $\cfgnt{r}_a < \cfgnt{r}_a^\prime$ based on a lexical
ordering of initialized references. Additionally
$\theta_\mathit{orig}$ allows for the possibility that the field
continues to remain uninitialized since the heap is symbolic. We refer
to this initialization in the summary machine as conditional
initialization. Finally, the fields of the reference $\cfgnt{r}_f$ are marked
as uninitialized. In the post condition of the rule the reference
$\cfgnt{r}_f$ points to $\theta$ which contains the union of all the
constraint location pairs sets, and the field $f$ points to the new
reference $\cfgnt{r}_f$ in the summary heap.

@@

We visualize the initialization process in~\figref{fig:initHeap}. The
graph in~\figref{fig:initHeap}(a) represents the initial heap. The
references with the superscript $s$ indicates that it is a local
reference. In~\figref{fig:initHeap} $\cfgnt{r}_0^s$ represents a local
reference for the $\cfgt{this}$ reference which has two fields $x$
and $y$ of the same type; reference $\cfgnt{r}_0^s$ points to location
$\cfgnt{l}_0$. Note that when no constraint is specified in the graphs, then
there is a implicit $\mathit{true}$ constraint, for example, $\cfgnt{r}_0^s$
points to $\cfgnt{l}_0$ on the constraint $\mathit{true}$. The fields $x$ and
$y$ point to the uninitialized reference $\cfgnt{r}_\mathit{un}$.

@@

The graph in~\figref{fig:initHeap}(b) represents the symbolic heap
summary after the initialization of the $\lp\cfgt{this}\  \cfgt{\$}\ \cfgnt{x}\rp$ field while
the graph in~\figref{fig:initHeap}(c) represents the symbolic heap
summary after the initialization of the $\lp\cfgt{this}\  \cfgt{\$}\ \cfgnt{y}\rp$ field
following the initialization of $\lp\cfgt{this}\  \cfgt{\$}\ \cfgnt{x}\rp$. The list
in~\figref{fig:initHeap}(d) represents the various sets constructed in
the summarize rule during the initialization of $\lp\cfgt{this}\  \cfgt{\$}\ \cfgnt{y}\rp$. We
also define values of constraints used as labels in the graph.

@@

Following the initialization to the $\lp\cfgt{this}\  \cfgt{\$}\ \cfgnt{x}\rp$, when we
initialize $\lp\cfgt{this}\  \cfgt{\$}\ \cfgnt{y}\rp$ the resulting summary heap is shown
in~\figref{fig:initHeap}(c). The set $\rho$ shown
in~\figref{fig:initHeap}(d) contains the potential aliases for
 $\lp\cfgt{this}\  \cfgt{\$}\ \cfgnt{y}\rp$ which is location $\cfgnt{l}_1$ ($\lp\cfgt{this}\  \cfgt{\$}\ \cfgnt{x}\rp$). The
constraint location $\theta$ sets represents the initialization
choices for $\lp\cfgt{this}\  \cfgt{\$}\ \cfgnt{y}\rp$. The new input reference $\cfgnt{r}_2^i$ for
$\lp\cfgt{this}\  \cfgt{\$}\ \cfgnt{y}\rp$ points to (a) location $\cfgnt{l}_\mathit{null}$ under the constraint
$\phi_{2a} := \cfgnt{r}_2^i = \cfgnt{r}_\mathit{null}$, (b) location $\cfgnt{l}_2$ under the
constraint $\phi_{2b} := \cfgnt{r}_2^i \neq \cfgnt{r}_\mathit{null} \wedge \cfgnt{r}_2^i \neq
\cfgnt{r}_1^i$ that states that $\cfgnt{r}_2^i$ is not null and does not alias
$\cfgnt{r}_1^i$, and (c) location $\cfgnt{l}_1$ under the constraint $\phi_{2c} :=
\cfgnt{r}_2^i \neq \cfgnt{r}_\mathit{null} \wedge \cfgnt{r}_2^i = \cfgnt{r}_1^i$ that states $\cfgnt{r}_2^i$
is not null and \emph{does} alias $\cfgnt{r}_1^i$.

>>>>

\section{Generating Summary Heaps}
\label{sec:precise}

In this section we present the rewrite rules for the $\rsym$ relation. 
There are three sets of rewrite rules specific to the summary
heap algorithm: (i) rules to initialize symbolic
references, (ii) field access and write rules, and (iii) rules
to check equality and inequality of references. These rules
form the core of the summary heap algorithm.

@@

During initialization of symbolic references in the $\rgse$ relation,
non-determinitic choices are introduced into the search to
account for initiaization of the reference to $\mathit{null}$, a new instance of the 
object, and to any type compatible object previously
initialized. In contrast, initialization rules for the $\rsym$ relation are 
defined on the bipartite graph consisting of references and locations,
and create a single branch containing a summarization of
all the initializations in a single graph.

@@

The initialization rules are invoked whenever an uninitialized field in a
symbolic reference is accessed. The function $\mathbb{UN}(\cfgnt{L},
\cfgnt{R}, \cfgnt{r}, \cfgnt{f}) = \{\cfgnt{l}\ ...\}$ returns
constraint-location pairs in which field $\cfgnt{f}$ is
uninitialized:
\[
\begin{array}{rcl}
\mathbb{UN}(\cfgnt{L}, \cfgnt{R}, \cfgnt{r}, \cfgnt{f}) & = &\{ \lp\phi\ \cfgnt{l}\rp \mid \lp \phi\ \cfgnt{l}\rp  \in \cfgnt{L}\lp \cfgnt{r}\rp  \wedge \\
& & \ \ \ \ \exists \phi^\prime \lp \lp \phi^\prime\ \cfgnt{l}_\mathit{un}\rp  \in \cfgnt{L}\lp \cfgnt{R}\lp l,\cfgnt{f}\rp\rp \wedge \\
& & \ \ \ \ \ \ \ \ \mathbb{S}\lp \phi \wedge \phi^\prime \rp\rp\}\\
\end{array}
\]
where $\mathbb{S}(\phi)$ returns true if $\phi$ is
satisfiable. Intutively, for reference $\cfgnt{r}$, this function constructs
the set, $\theta$ containing all contraint-location pairs that
point to field $\cfgnt{f}$ and $\cfgnt{f}$ points to
$\cfgnt{l}_\mathit{un}$.

@@

The rewrite rule to initialize a heap in the summary heap algorithm
is shown in~\figref{fig:symInit}. Given the unitialized set
$\mathbb{UN}$ for field $f$, the $\mathit{min}_l$ function returns
$(\phi_x\ \cfgnt{l}_x)$ based on a lexical ordering of uninitialized locations
in $\Lambda$. The init$_r$ method returns a new input reference $\cfgnt{r}_f$,
while the fresh$_f$ returns a new location of type $C$. The set $\rho$
contains the set of reference-location pairs, $(\cfgnt{r}_a\ \cfgnt{l}_a)$, the
field being initialized may alias. A pair $(\cfgnt{r}_a\ \cfgnt{l}_a)$ is contained
in the set $ \rho $ if the following conditions hold: (i) $\cfgnt{r}_a$ is
part of the input references initialized in the summary heap, (ii)
$\cfgnt{r}_a$ is in the pre-image of $\cfgnt{l}_a$, and (iii) the type 
of $\cfgnt{l}_a$ is the same as the type of the field being initialized.

There are three constraint-location pair sets that correspond to the
non-deterministic choices in GSE: (a) $\theta_\mathit{null}$ contains
a constraint $\cfgnt{r}_f = \cfgnt{r}_\mathit{null}$ paired with the location
$\cfgnt{l}_\mathit{null}$, (b) \sjp{something is not right in this choice}
$\theta_\mathit{new}$ creates a pair with a
constraint representing the fact that $\cfgnt{r}_f$ is not null and it does 
not alias any reference in set $\rho$ under and a newly created location 
$\cfgnt{l}_f$, and (c) 
$\theta_\mathit{alias}$ where for each element in the set $(\cfgnt{r}_a\ \cfgnt{l}_a)
\in \rho$ it creates a constraint stating that $\cfgnt{r}_f$ is not null, points to
$\cfgnt{r}_a$, and does not point to any reference contained in $(\cfgnt{r}_a^\prime\ 
\cfgnt{l}_a^\prime) \in \rho$ such that $\cfgnt{r}_a < \cfgnt{r}_a^\prime$ based on a lexical
ordering of initialized references. Additionally,
$\theta_\mathit{orig}$ allows for the possibility that the field
continues to remain uninitialized since the heap is symbolic.
\sjp{The last statement seems to jump out since we have not been
talking about the heap being symbolic (have we?)} We refer
to initialization in the summary machine as conditional
initialization \sjp{Not sure why we say this - what does it mean
to the reader?}. Finally, the fields of the reference $\cfgnt{r}_f$ are marked
as uninitialized. In the post condition of the rule, the reference
$\cfgnt{r}_f$ points to $\theta$ which contains the union of all the
constraint location pairs sets, and the field $f$ points to the new
reference $\cfgnt{r}_f$ in the summary heap.

@@

We visualize the initialization process in~\figref{fig:initHeap}. The
graph in~\figref{fig:initHeap}(a) represents the initial heap. The
references with the superscript $s$ indicates that it is a local
reference. In~\figref{fig:initHeap} $\cfgnt{r}_0^s$ represents a local
reference for the $\cfgt{this}$ reference which has two fields $x$
and $y$ of the same type; reference $\cfgnt{r}_0^s$ points to location
$\cfgnt{l}_0$. Note that when no constraint is specified in the graphs, then
there is an implicit $\mathit{true}$ constraint, for example, $\cfgnt{r}_0^s$
points to $\cfgnt{l}_0$ on the constraint $\mathit{true}$. The fields $x$ and
$y$ point to the uninitialized reference $\cfgnt{r}_\mathit{un}$.

@@

The graph in~\figref{fig:initHeap}(b) represents the summary heap
after the initialization of the $\lp\cfgt{this}\  \cfgt{\$}\ \cfgnt{x}\rp$ field while
the graph in~\figref{fig:initHeap}(c) represents the symbolic heap
summary after the initialization of the $\lp\cfgt{this}\  \cfgt{\$}\ \cfgnt{y}\rp$ field
following the initialization of $\lp\cfgt{this}\  \cfgt{\$}\ \cfgnt{x}\rp$. The list
in~\figref{fig:initHeap}(d) represents the various sets constructed in
the summarize rule during the initialization of $\lp\cfgt{this}\  \cfgt{\$}\ \cfgnt{y}\rp$. We
also define values of constraints used as labels in the graph.

@@

Following initialization to $\lp\cfgt{this}\  \cfgt{\$}\ \cfgnt{x}\rp$, when
$\lp\cfgt{this}\  \cfgt{\$}\ \cfgnt{y}\rp$ is initialized, the resulting summary heap is shown
in~\figref{fig:initHeap}(c). The set $\rho$ shown
in~\figref{fig:initHeap}(d) contains the potential aliases for
 $\lp\cfgt{this}\  \cfgt{\$}\ \cfgnt{y}\rp$ which is location $\cfgnt{l}_1$ ($\lp\cfgt{this}\  \cfgt{\$}\ \cfgnt{x}\rp$). The
constraint location $\theta$ sets represents the initialization
choices for $\lp\cfgt{this}\  \cfgt{\$}\ \cfgnt{y}\rp$. The new input reference $\cfgnt{r}_2^i$ for
$\lp\cfgt{this}\  \cfgt{\$}\ \cfgnt{y}\rp$ points to (a) location $\cfgnt{l}_\mathit{null}$ under the constraint
$\phi_{2a} := \cfgnt{r}_2^i = \cfgnt{r}_\mathit{null}$, (b) location $\cfgnt{l}_2$ under the
constraint $\phi_{2b} := \cfgnt{r}_2^i \neq \cfgnt{r}_\mathit{null} \wedge \cfgnt{r}_2^i \neq
\cfgnt{r}_1^i$ that states that $\cfgnt{r}_2^i$ is not null and does not alias
$\cfgnt{r}_1^i$, and (c) location $\cfgnt{l}_1$ under the constraint $\phi_{2c} :=
\cfgnt{r}_2^i \neq \cfgnt{r}_\mathit{null} \wedge \cfgnt{r}_2^i = \cfgnt{r}_1^i$ that states $\cfgnt{r}_2^i$
is not null and \emph{does} alias $\cfgnt{r}_1^i$.
